---
title: "Built environment and mortality risk from cardiovascular disease and diabetes in Medellin, Colombia: an ecological study"
subtitle: "Econometric analysis of the corrected manuscript submitted to Landscape and Urban Planning journal"
output: html_notebook
author: Jorge E. Patiño, Andy Hong, Juan C. Duque, Kazem Rahimi, Silvana Zapata, and Verónica Lopera
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

Quantitative analysis of the relationships between built-environment 
features and public health outcomes in Medellin, Colombia, at the neighborhood 
(barrio) level. International collaboration between the Oxford Martin School at
Oxford University (UK), RiSE Group at EAFIT University (CO), and the Medellin's
Health Office (CO).  

**Acknowledgements:**  
This work was completed with support from the PEAK Urban 
Programme, supported by 501 UKRI’s Global Challenge Research Fund, 
Grant Ref.: ES/P011055/1. Additionally, Kazem Rahimi is in
receipt of grants from the National Institute for Health Research 
(NIHR) Oxford Biomedical Research Centre (BRC), the Oxford Martin School (OMS), 
University of Oxford and the British Heart Foundation grant ref: PG/18/65/33872. 
The views expressed are those of the authors and not necessarily those of the OMS, 
the UK National Health Service (NHS), the NIHR or the
Department of Health and Social Care. The funding organizations had no role in design or
conduct of the study; collection, management, analysis, and interpretation of the data;
preparation, review, or approval of the manuscript; or decision to submit the manuscript for
publication.


# Datasets
**Spatial units of analysis:** inputs/data_MDE_health_built_env_geo.shp  
**Variables:** data_MDE_health_built_env.csv

# Computing environment
Load libraries:
```{r}
library(sp) # for creating neighborhoods contiguity matrix
library(sf) # for handling spatial units of analysis
library(tidyverse) # for plots and maps
library(MASS) # for robust regression
library(mctest) # for multicolinearity tests
library(stargazer) # for models and dataset summaries
library(jtools) # for single effects plots, used for quadratic variables
library(gridExtra) # for arranging several plots in single graph
library(spdep) # for spatial dependence tests
library(fields) # for calculating distances for distance contiguity matrix
library(tmap) # for mapping
library(corrplot) # for correlation matrix
```

Load datasets and join:
```{r}
df = read_csv("../inputs/data_MDE_health_built_env.cvs")
sf = st_read("../inputs/data_MDE_health_built_env_geo.shp")
dataset = left_join(sf, df, by = "CODIGO")
rm(df, sf)
```

Remove variable X1;
```{r}
dataset$X1 = NULL
```

# Mapping
## Mortality rates
```{r}
tm_shape(dataset) + 
  tm_fill(col = c("DCS_avg", "IHD_avg", "CVD_avg", "DM_avg"),
          title = c("", "", "", ""),
          style = "quantile",
          ) + 
  tm_facets(ncol = 4) +
  tm_scale_bar(breaks = c(0, 2), position = c("RIGHT", "BOTTOM")) +
  tm_layout(asp = 0.9,
            legend.position = c("LEFT", "TOP"),
            legend.text.size = 0.35,
            legend.width = 0.35,
            panel.show = TRUE,
            panel.labels = c("Diseases of the Circulatory System", 
                             "Ischemic Heart Disesae", 
                             "Cerebrovascular Disease", 
                             "Diabetes Mellitus"),
            panel.label.size = 0.5,
            frame = "gray50",
            )
tmap_save(filename = "../outputs/fig 1 mortality rates maps.png",
          height = 5,
          width = 16,
          units = "cm",
          dpi = 300)
```

## Urban form metrics
```{r}
tm_shape(dataset) + 
  tm_fill(col = c("int_dens",
                  "int_dens_2",
                  "slope_mean",
                  "NDVI_mean",
                  "lu_entropy_a",
                  "p_area_ind",
                  "sl_p_ind",
                  "POI_dens",
                  "POI_dens_2",
                  "pdens_2017",
                  "pdens_2_t",
                  "f_hig_SES"),
          title = c("", "", "", "", "", "", "", "", "", "", "", ""),
          style = "jenks",
          n = 5
          ) + 
  tm_scale_bar(breaks = c(0, 2), position = c("RIGHT", "BOTTOM")) +
  tm_layout(asp = 1.1,
            legend.position = c("LEFT", "TOP"),
            legend.text.size = 0.5,
            legend.width = 0.4,
            panel.show = TRUE,
            panel.labels = c("Intersection density",
                             "Int. dens. squared",
                             "Average slope",
                             "Average NDVI",
                             "Land use mix",
                             "% industrial land",
                             "% ind. land at neighbors",
                             "Density of amenities",
                             "Dens. of amenities squared",
                             "Population density",
                             "Pop. density squared",
                             "F. high SE"),
            panel.label.size = 0.85,
            frame = FALSE,
            )
```


# Econometrics
## Dataset summary:
```{r}
stargazer(st_drop_geometry(dataset), type = "text", digits = 2)
```

## Correlation matrix:
```{r}
variables = c("DCS_avg",
              "IHD_avg",
              "CVD_avg",
              "DM_avg",
              "int_dens", 
              "int_dens_2",
              "slope_mean",
              "NDVI_mean",
              "lu_entropy_a",
              "p_area_ind",
              "sl_p_ind",
              "POI_dens",
              "POI_dens_2",
              "pdens_2017",
              "pdens_2_t",
              "f_hig_SES")
x = st_drop_geometry(dataset)[, variables]
cor_pred = cor(x, method = "spearman")
corrplot.mixed(cor_pred, lower.col = "black", tl.pos = "lt",  tl.cex = 0.5,
               title = "Spearman correlation coefficients", cl.cex = 0.5,
               number.cex = 0.5, mar = c(0.5, 2, 3, 1))
```

## Linear models
Notes: Land use entropy adjusted with firm location data from DANE (see paper for a detailed explanation).
```{r}
DCS_ols = lm( #OLS
  DCS_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a + 
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    I(pdens_2017^2) + 
    f_hig_SES, 
  data = dataset)
DCS_rl = rlm( # robust regression
  DCS_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    I(pdens_2017^2) + 
    f_hig_SES, 
  data = dataset)
IHD_ols = lm(
  IHD_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    I(pdens_2017^2) + 
    f_hig_SES, 
  data = dataset)
IHD_rl = rlm(
  IHD_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    I(pdens_2017^2) + 
    f_hig_SES, 
  data = dataset)
CVD_ols = lm(
  CVD_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    I(pdens_2017^2) + 
    f_hig_SES, 
  data = dataset)
CVD_rl = rlm(
  CVD_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    I(pdens_2017^2) + 
    f_hig_SES, 
  data = dataset)
DM_ols = lm(
  DM_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    I(pdens_2017^2) + 
    f_hig_SES, 
  data = dataset)
DM_rl = rlm(
  DM_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    I(pdens_2017^2) + 
    f_hig_SES, 
  data = dataset)
```

## Multicolinearity Condition Number
Calculated on OLS models
```{r}
paste0("DCS: ", omcdiag(DCS_ols, Inter = F)$odiags[6])
paste0("IHD: ", omcdiag(IHD_ols, Inter = F)$odiags[6])
paste0("CVD: ", omcdiag(CVD_ols, Inter = F)$odiags[6])
paste0("DM: ", omcdiag(DM_ols, Inter = F)$odiags[6])
```

## Spatial dependence tests
### Create contiguity matrices
```{r}
# build contiguity matrix
Wr = poly2nb(as_Spatial(dataset), queen = FALSE) # Rook matrix
Wq = poly2nb(as_Spatial(dataset), queen = TRUE) # Queen matrix
# convert matrix to lists
listWr = nb2listw(Wr, style = "W", zero.policy = TRUE)
listWq = nb2listw(Wq, style = "W", zero.policy = TRUE)
# inverse distance matrix
mycoords <- coordinates(as_Spatial(dataset))
mydm <- rdist(mycoords)                  # computes distance in projected coordinates
for(i in 1:dim(mydm)[1]) {mydm[i,i] = 0} # renders exactly zero all diagonal elements
mydm[mydm > 1000] <- 0                   # all distances > 1000 miles are set to zero
mydm <- ifelse(mydm!=0, 1/mydm, mydm)    # inverting distances
listWid <- mat2listw(mydm, style="W")    # create a (normalized) listw object
```

### Lagrange Multiplier (LM) tests
Test run on OLS models using three different contiguity matrix specifications:  
- Rook: `listWr`.  
- Queen: `listWq`.  
- Inverse distance: `listWid`.  

#### DCS
```{r}
LM_Wr = lm.LMtests(DCS_ols, listWr, test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
LM_Wq = lm.LMtests(DCS_ols, listWq, test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
LM_Wid = lm.LMtests(DCS_ols, listWid, test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
summary(LM_Wr)
summary(LM_Wq)
summary(LM_Wid)
rm(LM_Wr, LM_Wq, LM_Wid)
```

#### IHD
```{r}
LM_Wr = lm.LMtests(IHD_ols, listWr, test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
LM_Wq = lm.LMtests(IHD_ols, listWq, test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
LM_Wid = lm.LMtests(IHD_ols, listWid, test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
summary(LM_Wr)
summary(LM_Wq)
summary(LM_Wid)
rm(LM_Wr, LM_Wq, LM_Wid)
```

#### CeVD
```{r}
LM_Wr = lm.LMtests(CVD_ols, listWr, test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
LM_Wq = lm.LMtests(CVD_ols, listWq, test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
LM_Wid = lm.LMtests(CVD_ols, listWid, test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
summary(LM_Wr)
summary(LM_Wq)
summary(LM_Wid)
rm(LM_Wr, LM_Wq, LM_Wid)
```

#### DM
```{r}
LM_Wr = lm.LMtests(DM_ols, listWr, test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
LM_Wq = lm.LMtests(DM_ols, listWq, test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
LM_Wid = lm.LMtests(DM_ols, listWid, test=c("LMerr", "LMlag", "RLMerr", "RLMlag"))
summary(LM_Wr)
summary(LM_Wq)
summary(LM_Wid)
rm(LM_Wr, LM_Wq, LM_Wid)
```

## Model's diagnostic plots
```{r}
par(mfrow = c(2,2), mar = c(4,4,5,2))
plot(DCS_ols, main = "DCS OLS")
plot(DCS_rl, main = "DCS Robust Linear")
plot(IHD_ols, main = "IHD OLS")
plot(IHD_rl, main = "IHD Robust Linear")
plot(CVD_ols, main = "CVD OLS")
plot(CVD_rl, main = "CVD Robust Linear")
plot(DM_ols, main = "DM OLS")
plot(DM_rl, main = "DM Robust Linear")
```


## Models' residuals plots
### DCS
#### OLS residuals and map:
```{r}
par(mfrow = c(1,2))
plot(resid(DCS_ols), main = "resid(DCS_ols)")
hist(resid(DCS_ols), breaks = 50, main = "resid(DCS_ols)")
# add residuals to data frame
dataset$res_DCS_ols = resid(DCS_ols)
# map
tm_shape(dataset) + 
  tm_fill(col = "res_DCS_ols", title = "DCS OLS residuals", 
          style = "sd", n = 7, palette = "PiYG", midpoint = NA) +
  tm_legend(legend.outside = TRUE, legend.outside.position = "right")
```
#### Robust linear regression residuals and map:
```{r}
par(mfrow = c(1,2))
plot(resid(DCS_rl), main = "resid(DCS_rl)")
hist(resid(DCS_rl), breaks = 50, main = "resid(DCS_rl)")
# add residuals to data frame
dataset$res_DCS_rl = resid(DCS_rl)
# map
tm_shape(dataset) + 
  tm_fill(col = "res_DCS_rl", title = "DCS Robust linear residuals", 
          style = "sd", n = 7, palette = "PiYG", midpoint = NA) +
  tm_legend(legend.outside = TRUE, legend.outside.position = "right")
```


### IHD
#### OLS residuals and map:
```{r}
par(mfrow = c(1,2))
plot(resid(IHD_ols), main = "resid(IHD_ols)")
hist(resid(IHD_ols), breaks = 50, main = "resid(IHD_ols)")
# add residuals to data frame
dataset$res_IHD_ols = resid(IHD_ols)
# map
tm_shape(dataset) + 
  tm_fill(col = "res_IHD_ols", title = "IHD OLS residuals", 
          style = "sd", n = 7, palette = "PiYG", midpoint = NA) +
  tm_legend(legend.outside = TRUE, legend.outside.position = "right")
```

#### Robust linear regression residuals and map:
```{r}
par(mfrow = c(1,2))
plot(resid(IHD_rl), main = "resid(IHD_rl)")
hist(resid(IHD_rl), breaks = 50, main = "resid(IHD_rl)")
# add residuals to data frame
dataset$res_IHD_rl = resid(IHD_rl)
# map
tm_shape(dataset) + 
  tm_fill(col = "res_IHD_rl", title = "IHD Robust linear residuals", 
          style = "sd", n = 7, palette = "PiYG", midpoint = NA) +
  tm_legend(legend.outside = TRUE, legend.outside.position = "right")
```

### CVD
#### OLS residuals and map:
```{r}
par(mfrow = c(1,2))
plot(resid(CVD_ols), main = "resid(CVD_ols)")
hist(resid(CVD_ols), breaks = 50, main = "resid(CVD_ols)")
# add residuals to data frame
dataset$res_CVD_ols = resid(CVD_ols)
# map
tm_shape(dataset) + 
  tm_fill(col = "res_CVD_ols", title = "CVD OLS residuals", 
          style = "sd", n = 7, palette = "PiYG", midpoint = NA) +
  tm_legend(legend.outside = TRUE, legend.outside.position = "right")
```

#### Robust linear regression residuals and map:
```{r}
par(mfrow = c(1,2))
plot(resid(CVD_rl), main = "resid(CVD_rl)")
hist(resid(CVD_rl), breaks = 50, main = "resid(CVD_rl)")
# add residuals to data frame
dataset$res_CVD_rl = resid(CVD_rl)
# map
tm_shape(dataset) + 
  tm_fill(col = "res_CVD_rl", title = "CVD Robust linear residuals", 
          style = "sd", n = 7, palette = "PiYG", midpoint = NA) +
  tm_legend(legend.outside = TRUE, legend.outside.position = "right")
```

### DM
#### OLS residuals and map:
```{r}
par(mfrow = c(1,2))
plot(resid(DM_ols), main = "resid(DM_ols)")
hist(resid(DM_ols), breaks = 50, main = "resid(DM_ols)")
# add residuals to data frame
dataset$res_DM_ols = resid(DM_ols)
# map
tm_shape(dataset) + 
  tm_fill(col = "res_DM_ols", title = "DM OLS residuals", 
          style = "sd", n = 7, palette = "PiYG", midpoint = NA) +
  tm_legend(legend.outside = TRUE, legend.outside.position = "right")
```

#### Robust linear regression residuals and map:
```{r}
par(mfrow = c(1,2))
plot(resid(DM_rl), main = "resid(DM_rl)")
hist(resid(DM_rl), breaks = 50, main = "resid(DM_rl)")
# add residuals to data frame
dataset$res_DM_rl = resid(DM_rl)
# map
tm_shape(dataset) + 
  tm_fill(col = "res_DM_rl", title = "DM Robust linear residuals", 
          style = "sd", n = 7, palette = "PiYG", midpoint = NA) +
  tm_legend(legend.outside = TRUE, legend.outside.position = "right")
```

## Effects plots
### Plots with local references
Values from 3 Medellin administrative neighborhoods.
#### Selected neighborhoods:  
- El Tesoro: int_dens 52, POI_dens 5.1, pdens_2017 8966  
- Sevilla: int_dens 207, POI_dens 118, pdens_2017 11046  
- Francisco Antonio Zea: int_dens 438, POI_dens 34, pdens_2017 61024  

#### Intersection density
Models with significant term: DCS, CVD, and DM
```{r}
p1 = effect_plot(DCS_rl, pred = int_dens, interval = TRUE, robust = TRUE) + 
  xlim(0, 500) + 
  #ylim(0, 150) +
  xlab("") +
  ylab("Diseases of the Circulatory System") + 
  geom_vline(xintercept = 52, colour = "orange", alpha = 0.5) + 
  geom_vline(xintercept = 207, colour = "orange", alpha = 0.5) + 
  geom_vline(xintercept = 438, colour = "orange", alpha = 0.5)
p2 = effect_plot(CVD_rl, pred = int_dens, interval = TRUE, robust = TRUE) +
  xlim(0, 500) + 
  #ylim(0, 30) +
  xlab("") +
  ylab("Cerebrovascular Disease") + 
  geom_vline(xintercept = 52, colour = "orange", alpha = 0.5) + 
  geom_vline(xintercept = 207, colour = "orange", alpha = 0.5) + 
  geom_vline(xintercept = 438, colour = "orange", alpha = 0.5)
p3 = effect_plot(DM_rl, pred = int_dens, interval = TRUE, robust = TRUE) +
  xlim(0, 500) +
  #ylim(0, 15) +
  xlab("Intersection density") +
  ylab("Diabetes Mellitus") + 
  geom_vline(xintercept = 52, colour = "orange", alpha = 0.5) +  
  annotate("text", 
           label = "El Tesoro", 
           x = 52, 
           y = 1, 
           size = 3, 
           colour = "orange", 
           angle = 90, 
           vjust = -0.15) + 
  geom_vline(xintercept = 207, colour = "orange", alpha = 0.5) + 
  annotate("text", 
           label = "Sevilla", 
           x = 207, 
           y = 1, 
           size = 3, 
           colour = "orange", 
           angle = 90, 
           vjust = -0.15) +
  geom_vline(xintercept = 438, colour = "orange", alpha = 0.5) + 
  annotate("text", 
           label = "F. A. Zea", 
           x = 438, 
           y = 1, 
           size = 3, 
           colour = "orange", 
           angle = 90, 
           vjust = -0.15)
g = arrangeGrob(p1, p2, p3, nrow = 3)
ggsave("../outputs/effects_intersection_density.png", g, height = 10, width = 2.5)
rm(p1, p2, p3, g)
```

#### Density of amenities
Models with significant term: DCS, IHD, and CVD
```{r}
p1 = effect_plot(DCS_rl, pred = POI_dens, interval = TRUE, robust = TRUE) +
  xlim(0, 300) +
  xlab("") +
  ylab("Diseases of the Circulatory System") +
  geom_vline(xintercept = 5, colour = "orange", alpha = 0.5) + 
  geom_vline(xintercept = 34, colour = "orange", alpha = 0.5) + 
  geom_vline(xintercept = 118, colour = "orange", alpha = 0.5)
p2 = effect_plot(IHD_rl, pred = POI_dens, interval = TRUE, robust = TRUE) +
  xlim(0, 300) +
  xlab("") +
  ylab("Ischemic Heart Disease") +
  geom_vline(xintercept = 5, colour = "orange", alpha = 0.5) + 
  geom_vline(xintercept = 34, colour = "orange", alpha = 0.5) + 
  geom_vline(xintercept = 118, colour = "orange", alpha = 0.5)
p3 = effect_plot(CVD_rl, pred = POI_dens, interval = TRUE, robust = TRUE) +
  xlim(0, 300) +
  xlab("Density of amenities") +
  ylab("Cerebrovascular Disease") +
  geom_vline(xintercept = 5, colour = "orange", alpha = 0.5) + 
  annotate("text", 
           label = "El Tesoro", 
           x = 5, 
           y = -50, 
           size = 3, 
           colour = "orange", 
           angle = 90, 
           vjust = -0.15) +
  geom_vline(xintercept = 34, colour = "orange", alpha = 0.5) + 
  annotate("text", 
           label = "F. A. Zea", 
           x = 34, 
           y = -50, 
           size = 3, 
           colour = "orange", 
           angle = 90, 
           vjust = -0.15) +
  geom_vline(xintercept = 118, colour = "orange", alpha = 0.5) + 
  annotate("text", 
           label = "Sevilla", 
           x = 118, 
           y = -50, 
           size = 3, 
           colour = "orange", 
           angle = 90, 
           vjust = -0.15)
g = arrangeGrob(p1, p2, p3, nrow = 3)
ggsave("../outputs/effects_dens_amenities.png", g, height = 10, width = 2.5)
rm(p1, p2, p3, g)
```

#### Population density
Plot for DCS, IHD, and CVD:
```{r}
p1 = effect_plot(DCS_rl, pred = pdens_2017, interval = TRUE, robust = TRUE) + 
  xlim(0, 80000) +
  xlab("") +
  ylab("Diseases of the Circulatory System") +
  geom_vline(xintercept = 8966, colour = "orange", alpha = 0.5) + 
  geom_vline(xintercept = 11046, colour = "orange", alpha = 0.5) + 
  geom_vline(xintercept = 61024, colour = "orange", alpha = 0.5)
p2 = effect_plot(IHD_rl, pred = pdens_2017, interval = TRUE, robust = TRUE) +
  xlim(0, 80000) +
  xlab("") +
  ylab("Ischemic Heart Disease") +
  geom_vline(xintercept = 8966, colour = "orange", alpha = 0.5) + 
  geom_vline(xintercept = 11046, colour = "orange", alpha = 0.5) + 
  geom_vline(xintercept = 61024, colour = "orange", alpha = 0.5)
p3 = effect_plot(CVD_rl, pred = pdens_2017, interval = TRUE, robust = TRUE) +
  xlim(0, 80000) +
  xlab("Population density") +
  ylab("Cerebrovascular Disease") +geom_vline(xintercept = 8966, colour = "orange", alpha = 0.5) + 
  annotate("text", 
           label = "El Tesoro", 
           x = 8966, 
           y = 10, 
           size = 3, 
           colour = "orange", 
           angle = 90, 
           vjust = -0.3) + 
  geom_vline(xintercept = 11046, colour = "orange", alpha = 0.5) +  
  annotate("text", 
           label = "Sevilla", 
           x = 11046, 
           y = 10, 
           size = 3, 
           colour = "orange", 
           angle = 90, 
           vjust = 1.2) +
  geom_vline(xintercept = 61024, colour = "orange", alpha = 0.5) + 
  annotate("text", 
           label = "F. A. Zea", 
           x = 61024, 
           y = 10, 
           size = 3, 
           colour = "orange", 
           angle = 90, 
           vjust = -0.4)
g = arrangeGrob(p1, p2, p3, nrow = 3)
ggsave("../outputs/effects_pop_density.png", g, height = 10, width = 2.5)
rm(p1, p2, p3, g)
```

## Models summary
### Rewrite nice models
Rewrite models with squared population density divided by 1,000,000:
```{r}
DCS_ols = lm( #OLS
  DCS_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a + 
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    pdens_2_t + 
    f_hig_SES, 
  data = dataset)
DCS_rl = rlm( # robust regression
  DCS_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    pdens_2_t + 
    f_hig_SES, 
  data = dataset)
IHD_ols = lm(
  IHD_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    pdens_2_t + 
    f_hig_SES, 
  data = dataset)
IHD_rl = rlm(
  IHD_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    pdens_2_t + 
    f_hig_SES, 
  data = dataset)
CVD_ols = lm(
  CVD_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    pdens_2_t + 
    f_hig_SES, 
  data = dataset)
CVD_rl = rlm(
  CVD_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    pdens_2_t + 
    f_hig_SES, 
  data = dataset)
DM_ols = lm(
  DM_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    pdens_2_t + 
    f_hig_SES, 
  data = dataset)
DM_rl = rlm(
  DM_avg ~ 
    int_dens + 
    I(int_dens^2) + 
    slope_mean + 
    NDVI_mean + 
    lu_entropy_a +
    p_area_ind + 
    sl_p_ind + 
    POI_dens + 
    I(POI_dens^2) + 
    pdens_2017 +
    pdens_2_t + 
    f_hig_SES, 
  data = dataset)
```

### Summary
Summary table:
```{r}
stargazer(DCS_ols, DCS_rl, IHD_ols, IHD_rl, CVD_ols, CVD_rl, DM_ols, DM_rl,
          keep.stat = c("rsq", "adj.rsq", "ser", "f"),
          dep.var.labels = c("DCS", "IHD", "CeVD", "DM"),
          covariate.labels = c("Constant", 
                               "Intersection density",
                               "Int. dens. squared",
                               "Average slope",
                               "Average NDVI",
                               "Land use mix",
                               "% industrial land",
                               "% ind. land at neighbors",
                               "Density of amenities",
                               "Dens. of amenities squared",
                               "Population density",
                               "Pop. density squared",
                               "F. high SE"),
          type = "text", align = TRUE, no.space = FALSE, intercept.bottom = FALSE, intercept.top = TRUE, df = FALSE,
          title = "Health and built environment relationships in Medellin, Colombia",
          add.lines = list(c("Condition Number", "23.4", "", "23.4", "", "23.4", "", "23.4", ""),
                           c("", "", "", "", "", "", "", "", ""),
                           c("Spatial dependence tests", "", "", "", "", "", "", "", ""),
                           c("LM Error", "2.1149", "", "0.0025", "", "0.3105", "", "0.5479", ""),
                           c("LM Lag", "1.4287", "", "1.2549", "", "1.3569", "", " 1.1056", ""),
                           c("", "", "", "", "", "", "", "", "")),
          notes = c("Pop. density squared divided by 1,000,000",
                    "N = 245"))
```




END.